# pre-class


Make sure you commit this often with meaningfull messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```{r}
install.packages("gapminder")
```



```{r}
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
gapminder %>%
  
  #group dacta into continents
  group_by(continent) %>%
  
  #select only country and continent columns
  select(country, continent) %>%
  
  #remove duplicate rows
  distinct() %>%
  
  #count number of countries in each group/continent
  summarize("Unique Countries"= n())
```

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
gapminder %>%
  
  #return rows with data from 1997
  filter(continent == "Europe", year == 1997) %>%
  
  #display data arranged by increasing GDP per capita
  arrange(gdpPercap)
```

In 1997, the European nation with the lowest GDP per capita was Albania which had a GDP of 3193.055 per capita.

3. According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
gapminder %>%
  
  #group data by continent
  group_by(continent) %>%
  
  #return rows containing data from the 80s
  filter(year >= 1980 & year < 1990) %>%
  
  #Report average life expectancy for each group/continent
  summarize("Average Life Expectancy" = mean(lifeExp))


```


4. What 5 countries have the highest total GDP over all years combined?

```{r}
gapminder %>%
  
  #group dataset by country
  group_by(country) %>%
  
  #calculate total GDP, multiplying by population to convert GDP per capita values
  #Store result in new dataframe, Totals, so we can display the results in descending order
  summarize("Total_GDP" = sum(gdpPercap * pop)) -> Totals

Totals %>%
  #display results by descenging Total GDP
  arrange(desc(Total_GDP))
```

The 5 countries with the highest total GDP over all years combined are, in decreasing order: the United States, Japan, China, Germany, and the United Kingdom.

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._
6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).

